---
title: "model_calibration"
output: html_document
date: "2023-10-30"
---

### Load required packages

```{r}
pacman::p_load(tidyverse, tidymodels, ggplot2, patchwork, rpart, MASS, dplyr, poissonreg, MASS, parsnip, dplyr)
```

### Read data

```{r}
data = read_csv("data/april_joinedv2.csv")
# head(data, 5)
names(data)
```

### Splitting data set for different scenarios

```{r}
# from home and bus stop to office, school and bus stop
weekday_morning <- data %>%
  filter(DAY_TYPE == "WEEKDAY" & TIME_PER_HOUR %in% c(6, 9)) %>%
  dplyr::select(TOTAL_TRIPS,
         origin_HDB, 
         origin_HDB_Dus,
         origin_bus, 
         destination_Business, 
         destination_FinInsts,
         destination_bus,
         origin_station,
         destination_station) %>%
  mutate_at(vars(starts_with("origin_"), starts_with("destination_")), ~ifelse(is.na(.), 0, .)) %>%
  group_by(origin_station, destination_station) %>%
  mutate(sum())

#from office, school and bus stop to home, bus stops and pubs
weekday_evening <- data %>%
  filter(DAY_TYPE == "WEEKDAY" & TIME_PER_HOUR %in% c(5, 8)) %>%
  dplyr::select(TOTAL_TRIPS,
         origin_Business, 
         origin_FinInsts,
         origin_bus, 
         destination_HDB, 
         destination_HDB_Dus,
         destination_bus,
         destination_Pubs,
         origin_station,
         destination_station) %>%
  mutate_at(vars(starts_with("origin_"), starts_with("destination_")), ~ifelse(is.na(.), 0, .)) %>%
  group_by(origin_station, destination_station) %>%
  mutate(sum())

#from home, bus stop to leisure
weekend_morning <- data %>%
  filter(DAY_TYPE == "WEEKENDS/HOLIDAY" & TIME_PER_HOUR %in% c(11, 14)) %>%
  dplyr::select(TOTAL_TRIPS,
         origin_HDB, 
         origin_HDB_Dus,
         origin_bus, 
         destination_Cinema, 
         destination_ParkRec,
         destination_Malls,
         destination_RetailShop,
         destination_bus,
         origin_station,
         destination_station) %>%
  mutate_at(vars(starts_with("origin_"), starts_with("destination_")), ~ifelse(is.na(.), 0, .)) %>%
  group_by(origin_station, destination_station) %>%
  mutate(sum())

#from leisure to home and bus stop
weekend_evening <- data %>%
  filter(DAY_TYPE == "WEEKENDS/HOLIDAY" & TIME_PER_HOUR %in% c(16, 19)) %>%
  dplyr::select(TOTAL_TRIPS,
         origin_Cinema, 
         origin_ParkRec,
         origin_Malls,
         origin_RetailShop,
         origin_bus,
         destination_HDB,
         destination_HDB_Dus,
         destination_bus,
         origin_station,
         destination_station) %>%
  mutate_at(vars(starts_with("origin_"), starts_with("destination_")), ~ifelse(is.na(.), 0, .)) %>%
  group_by(origin_station, destination_station) %>%
  mutate(sum())
```

### Visualizing the relationships in filtered data set

```{r}
# y_vars <- names(weekend_morning)[2:13]
# plot_list <- list()
# 
# for (y in y_vars) {
#   # Create the plot
#   plot_title <- paste(y)
#   p <- ggplot(weekend_morning, aes_string(x = "TOTAL_TRIPS", y = y)) +
#     geom_point(size=1) +
#     ggtitle(plot_title) +
#     xlab("total_trips")
#   
#   # Add the plot to the list
#   plot_list[[y]] <- p
# }
# 
# # Combine all plots in the list into a single plot
# combined_plot <- wrap_plots(plotlist = plot_list)
# 
# # Print the combined plot
# print(combined_plot)
```

### Splitting data - sampling and cross-validation

```{r}
# Split data into train, validation, and test sets using rsample
set.seed(1243)
# 80% data for training, 20% for testing
wdm <- initial_split(weekday_morning, prop = 0.8, strata = "TOTAL_TRIPS")
wdm_train <- training(wdm)
wdm_test <- testing(wdm)

wde <- initial_split(weekday_evening, prop = 0.8, strata = "TOTAL_TRIPS")
wde_train <- training(wde)
wde_test <- testing(wde)

wem <- initial_split(weekend_morning, prop = 0.8, strata = "TOTAL_TRIPS")
wem_train <- training(wem)
wem_test <- testing(wem)

wee <- initial_split(weekend_evening, prop = 0.8, strata = "TOTAL_TRIPS")
wee_train <- training(wee)
wee_test <- testing(wee)
```

### Data pre-processing with recipe

```{r}
wdm_recipe <- recipe(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, data = wdm_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  # step_log(all_numeric()) %>%
  prep() %>%
  juice()

wde_recipe <- recipe(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, data = wde_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  # step_log(all_numeric()) %>%
  prep() %>%
  juice()

wem_recipe <- recipe(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, data = wem_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  # step_log(all_numeric()) %>%
  prep() %>%
  juice()

wee_recipe <- recipe(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, data = wee_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  # step_log(all_numeric()) %>%
  prep() %>%
  juice()
```

### [STAT] Multiple Linear Regression - lm engine

```{r}
wdm_lm_fit <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("glm") %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, wdm_recipe)

wde_lm_fit <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("glm") %>%
  fit(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, wde_recipe)

wem_lm_fit <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("glm") %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, wem_recipe)

wee_lm_fit <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("glm") %>%
  fit(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, wee_recipe)
```

### [STAT] Poisson - glm engine
```{r}
wdm_poisson_fit <- poisson_reg() %>%
  set_engine("glm") %>%
  set_mode("regression") %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, data = wdm_recipe)

wde_poisson_fit <- poisson_reg() %>%
  set_engine("glm") %>%
  set_mode("regression") %>%
  fit(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, data = wde_recipe)

wem_poisson_fit <- poisson_reg() %>%
  set_engine("glm") %>%
  set_mode("regression") %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, data = wem_recipe)

wee_poisson_fit <- poisson_reg() %>%
  set_engine("glm") %>%
  set_mode("regression") %>%
  fit(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, data = wee_recipe)

```

### [STAT] Negative Binomial - glm engine (corrupt)
```{r}
wdm_nm_fit <- glm.nb(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, data = wdm_recipe)

wde_nm_fit <- glm.nb(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, data = wde_recipe)

wem_nm_fit <- glm.nb(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, data = wem_recipe)

wee_nm_fit <- glm.nb(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, data = wee_recipe)

```

### [ML] Regression Tree - rpart (what we want from this?)
```{r}
wdm_rt_fit <- rpart(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, data = wdm_recipe, method="poisson")

wde_rt_fit <- rpart(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, data = wde_recipe)

wem_rt_fit <- rpart(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, data = wem_recipe)

wee_rt_fit <- rpart(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, data = wee_recipe)

plot(wdm_rt_fit)
```

### [ML] Random Forest
```{r}
wdm_rf_fit <- rand_forest(mode = "regression", engine = "randomForest", mtry = .preds(), trees = 100) %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, wdm_recipe)

wde_rf_fit <- rand_forest(mode = "regression", engine = "randomForest", mtry = .preds(), trees = 100) %>%
  fit(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, wde_recipe)

wem_rf_fit <- rand_forest(mode = "regression", engine = "randomForest", mtry = .preds(), trees = 100) %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, wem_recipe)

wee_rf_fit <- rand_forest(mode = "regression", engine = "randomForest", mtry = .preds(), trees = 100) %>%
  fit(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, wee_recipe)

```

### [ML] Boost Tree - xgboost
```{r}
wdm_bt_fit <- boost_tree() %>%
  set_mode("regression") %>%
  set_engine("xgboost") %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, wdm_recipe)
  
wde_bt_fit <- boost_tree() %>%
  set_mode("regression") %>%
  set_engine("xgboost") %>%
  fit(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, wde_recipe)
  
wem_bt_fit <- boost_tree() %>%
  set_mode("regression") %>%
  set_engine("xgboost") %>%
  fit(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, wem_recipe)
 
wee_bt_fit <- boost_tree() %>%
  set_mode("regression") %>%
  set_engine("xgboost") %>%
 fit(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, wee_recipe)

```

### Computing prediction values
```{r}
wdm_trip_recipe <- recipe(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Business + destination_FinInsts + destination_bus, data=wdm_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_dummy(all_nominal()) %>%
  step_log(all_outcomes()) %>%
  step_scale(all_numeric()) %>%
  prep()

wde_trip_recipe <- recipe(TOTAL_TRIPS ~ origin_Business + origin_FinInsts + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus + destination_Pubs, data = wde_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_dummy(all_nominal()) %>%
  prep()

wem_trip_recipe <- recipe(TOTAL_TRIPS ~ origin_HDB + origin_HDB_Dus + origin_bus + destination_Cinema + destination_ParkRec + destination_Malls + destination_RetailShop + destination_bus, data = wem_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_dummy(all_nominal()) %>%
  prep()

wee_trip_recipe <- recipe(TOTAL_TRIPS ~ origin_Cinema + origin_ParkRec + origin_Malls + origin_RetailShop + origin_bus + destination_HDB + destination_HDB_Dus + destination_bus, data = wee_train) %>%
  step_normalize(all_predictors(), -all_nominal()) %>%
  step_dummy(all_nominal()) %>%
  prep()

wdm_bake <- bake(wdm_trip_recipe, new_data = wdm_test, all_predictors())
wde_bake <- bake(wde_trip_recipe, new_data = wde_test, all_predictors())
wem_bake <- bake(wem_trip_recipe, new_data = wem_test, all_predictors())
wee_bake <- bake(wee_trip_recipe, new_data = wee_test, all_predictors())

```

### LM predictions
```{r}
wdm_lm_pred <- predict(wdm_lm_fit, wdm_bake) %>%
    rename(wdm_lm = .pred)

wde_lm_pred <- predict(wde_lm_fit, wde_bake) %>%
    rename(wde_lm = .pred)

wem_lm_pred <- predict(wem_lm_fit, wem_bake) %>%
    rename(wem_lm = .pred)

wee_lm_pred <- predict(wee_lm_fit, wee_bake) %>%
    rename(wee_lm = .pred)

```

### Poisson predictions
```{r}
wdm_poisson_pred <- predict(wdm_poisson_fit, wdm_bake) %>%
    rename(wdm_poisson = .pred)

wde_poisson_pred <- predict(wde_poisson_fit, wde_bake) %>%
    rename(wde_poisson = .pred)

wem_poisson_pred <- predict(wem_poisson_fit, wem_bake) %>%
    rename(wem_poisson = .pred)

wee_poisson_pred <- predict(wee_poisson_fit, wee_bake) %>%
    rename(wee_poisson = .pred)
```

### NM predictions
```{r}
wdm_nm_pred <- predict(wdm_nm_fit, wdm_bake)
# %>%
#     rename(wdm_nm = .pred)

wde_nm_pred <- predict(wde_nm_fit, wde_bake)

wem_nm_pred <- predict(wem_nm_fit, wem_bake)

wee_nm_pred <- predict(wee_nm_fit, wee_bake)
```

### RT predictions
```{r}
# rpart is tree
```

### RF predictions
```{r}
wdm_rf_pred <- predict(wdm_rf_fit, wdm_bake) %>%
    rename(wdm_rf = .pred)

wde_rf_pred <- predict(wde_rf_fit, wde_bake) %>%
    rename(wde_rf = .pred)

wem_rf_pred <- predict(wem_rf_fit, wem_bake) %>%
    rename(wem_rf = .pred)

wee_rf_pred <- predict(wee_rf_fit, wee_bake) %>%
    rename(wee_rf = .pred)
```

### BT predictions
```{r}
wdm_bt_pred <- predict(wdm_bt_fit, wdm_bake) %>%
    rename(wdm_bt = .pred)

wde_bt_pred <- predict(wde_bt_fit, wde_bake) %>%
    rename(wde_bt = .pred)

wem_bt_pred <- predict(wem_bt_fit, wem_bake) %>%
    rename(wem_bt = .pred)

wee_bt_pred <- predict(wee_bt_fit, wee_bake) %>%
    rename(wee_bt = .pred)
```

### Model Comparison (continue when backs)
```{r}
wdm_test_results <- wdm_test %>%
  dplyr::select(TOTAL_TRIPS) %>%
  bind_cols(wdm_lm_pred) %>%
  bind_cols(wdm_poisson_pred) %>%
  bind_cols(wdm_nm_pred) %>%
  bind_cols(wdm_rf_pred) %>%
  bind_cols(wdm_bt_pred)

wde_test_results <- wde_test %>%
  dplyr::select(TOTAL_TRIPS) %>%
  bind_cols(wde_lm_pred) %>%
  bind_cols(wde_poisson_pred) %>%
  bind_cols(wde_nm_pred) %>%
  bind_cols(wde_rf_pred) %>%
  bind_cols(wde_bt_pred)

wem_test_results <- wem_test %>%
  dplyr::select(TOTAL_TRIPS) %>%
  bind_cols(wem_lm_pred) %>%
  bind_cols(wem_poisson_pred) %>%
  bind_cols(wem_nm_pred) %>%
  bind_cols(wem_rf_pred) %>%
  bind_cols(wem_bt_pred)

wee_test_results <- wee_test %>%
  dplyr::select(TOTAL_TRIPS) %>%
  bind_cols(wee_lm_pred) %>%
  bind_cols(wee_poisson_pred) %>%
  bind_cols(wee_nm_pred) %>%
  bind_cols(wee_rf_pred) %>%
  bind_cols(wee_bt_pred)
```

```{r}
metric_set(rmse, rsq, mase)
```

### Visualising performance
```{r}
wdm_test_results %>%
  ungroup() %>%
  rename(wdm_nb = "...6") %>%
  dplyr::select(c(3:8)) %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  group_by(model) %>%
  metrics(truth = TOTAL_TRIPS, estimate = prediction) 

wde_test_results %>%
  ungroup() %>%
  rename(wdm_nb = "...6") %>%
  dplyr::select(c(3:8)) %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  group_by(model) %>%
  metrics(truth = TOTAL_TRIPS, estimate = prediction)

wem_test_results %>%
  ungroup() %>%
  rename(wdm_nb = "...6") %>%
  dplyr::select(c(3:8)) %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  group_by(model) %>%
  metrics(truth = TOTAL_TRIPS, estimate = prediction)

wee_test_results %>%
  ungroup() %>%
  rename(wdm_nb = "...6") %>%
  dplyr::select(c(3:8)) %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  group_by(model) %>%
  metrics(truth = TOTAL_TRIPS, estimate = prediction)
```

```{r}
#use fig.width and fig.height to edit size 
wdm_test_results %>%
  ungroup() %>%
  dplyr::select(c(3:8)) %>%
  rename(wdm_nb = "...6") %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  ggplot(aes(x = prediction, y = TOTAL_TRIPS)) + 
  geom_abline(col = "red", lty = 2) +
  geom_point(alpha = .4) +
  facet_wrap(~model) +
  coord_fixed() +
  ggtitle("Weekday morning model results")

wde_test_results %>%
  ungroup() %>%
  dplyr::select(c(3:8)) %>%
  rename(wde_nb = "...6") %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  ggplot(aes(x = prediction, y = TOTAL_TRIPS)) + 
  geom_abline(col = "red", lty = 2) +
  geom_point(alpha = .4) +
  facet_wrap(~model) +
  coord_fixed() +
  ggtitle("Weekday evening model results")

wem_test_results %>%
  ungroup() %>%
  dplyr::select(c(3:8)) %>%
  rename(wem_nb = "...6") %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  ggplot(aes(x = prediction, y = TOTAL_TRIPS)) + 
  geom_abline(col = "red", lty = 2) +
  geom_point(alpha = .4) +
  facet_wrap(~model) +
  coord_fixed() +
  ggtitle("Weekend morning model results")

wee_test_results %>%
  ungroup() %>%
  dplyr::select(c(3:8)) %>%
  rename(wee_nb = "...6") %>%
  pivot_longer(!TOTAL_TRIPS, names_to = "model", values_to = "prediction") %>%
  ggplot(aes(x = prediction, y = TOTAL_TRIPS)) + 
  geom_abline(col = "red", lty = 2) +
  geom_point(alpha = .4) +
  facet_wrap(~model) +
  coord_fixed() +
  ggtitle("Weekend evening model results")
```